---
title: "Biostat 203B Homework 1 Solution"
author: Siyan (Clare) Chen
subtitle: Due Jan 22 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Display machine information for reproducibility:
```{r, eval=F}
sessionInfo()
```

## Q1. Git/GitHub

**No handwritten homework reports are accepted for this course.** We work with Git and GitHub. Efficient and abundant use of Git, e.g., frequent and well-documented commits, is an important criterion for grading your homework.

1. Apply for the [Student Developer Pack](https://education.github.com/pack) at GitHub using your UCLA email.

**Solution:** Done

2. Create a **private** repository `biostat-203b-2021-winter` and add `Hua-Zhou`, `Chris-German` and `ElvisCuiHan` as your collaborators with write permission.

**Solution:** Done

3. Top directories of the repository should be `hw1`, `hw2`, ... Maintain two branches `master` and `develop`. The `develop` branch will be your main playground, the place where you develop solution (code) to homework problems and write up report. The `master` branch will be your presentation area. Submit your homework files (R markdown file `Rmd`, `html` file converted from R markdown, all code and data sets to reproduce results) in `master` branch.

**Solution:** Done

4. After each homework due date, teaching assistant and instructor will check out your master branch for grading. Tag each of your homework submissions with tag names `hw1`, `hw2`, ... Tagging time will be used as your submission time. That means if you tag your `hw1` submission after deadline, penalty points will be deducted for late submission.

**Solution:** Done

5. After this course, you can make this repository public and use it to demonstrate your skill sets on job market.

**Solution:** Nice :)

## Q2. Linux Shell Commands

1. This exercise (and later in this course) uses the [MIMIC-IV data](https://mimic-iv.mit.edu), a freely accessible critical care database developed by the MIT Lab for Computational Physiology. Follow the instructions at <https://mimic-iv.mit.edu/docs/access/> to (1) complete the CITI `Data or Specimens Only Research` course and (2) obtain the PhysioNet credential for using the MIMIC-IV data. Display the verification links to your completion report and completion certificate here. (Hint: The CITI training takes a couple hours and the PhysioNet credentialing takes a couple days; do not leave it to the last minute.)

**Solution:** 
Here is my [completion report](https://www.citiprogram.org/verify/?k796dd661-cb1d-4414-8030-36f2df8a133e-40357427).
Here is my [completion certificate](https://www.citiprogram.org/verify/?wbfa24715-0661-4cbf-a999-7a7d5e54bf68-40357427).

2. The `/usr/203b-data/mimic-iv/` folder on teaching server contains data sets from MIMIC-IV. Refer to <https://mimic-iv.mit.edu/docs/datasets/> for details of data files.  
    ```{bash}
    ls -l /usr/203b-data/mimic-iv
    ```
Please, do **not** put these data files into Git; they are big. Do **not** copy them into your directory. Do **not** decompress the gz data files. These create unnecessary big files on storage and are not big data friendly practices. Just read from the data folder `/usr/203b-data/mimic-iv` directly in following exercises. 

**Solution:** Got it

    Use Bash commands to answer following questions.

3. Display the contents in the folders `core`, `hosp`, `icu`. What are the functionalities of the bash commands `zcat`, `zless`, `zmore`, and `zgrep`? 

**Solution:**
    ```{bash}
    ls -l /usr/203b-data/mimic-iv/core
    ls -l /usr/203b-data/mimic-iv/hosp
    ls -l /usr/203b-data/mimic-iv/icu
    ```

`zcat` can be used to display the content of compressed file without uncompressing them in advance. 

`zmore` is used to browse a compressed file downwards by pages. We can view the next page by spacebar and exit by q buttom.

`zless` allows us to view the file both upwards and downwards by using different input. 

`zgrep` is used to search certain string in the compressed file and return the corresponding lines.

Overall, the functionalities of those commands are similar to `zcat`,`zmore`,`zless` and `zgrep`, but they are executed on compressed files instead.

4. What's the output of following bash script?
    ```{bash, eval=F}
    for datafile in /usr/203b-data/mimic-iv/core/*.gz
      do
        ls -l $datafile
      done
    ```
**Solution:** The script is a for loop that displays the .gz file in the assigned working directory.

Display the number of lines in each data file using a similar loop.
**Solution:**
    ```{bash}
    for datafile in /usr/203b-data/mimic-iv/core/*.gz
          do
           zcat $datafile | wc -l 
    done 
    ```


4. Display the first few lines of `admissions.csv.gz`. How many rows are in this data file? How many unique patients (identified by `subject_id`) are in this data file? What are the possible values taken by each of the variable `admission_type`, `admission_location`, `insurance`, `language`, `marital_status`, and `ethnicity`? Also report the count for each unique value of these variables. (Hint: combine Linux commands `zcat`, `head`/`tail`, `awk`, `uniq`, `wc`, and so on.)

**Solution:** See below the first 10 lines of `admissions.csv.gz`.
The total number of rows in this file is 65080 (found in the previous part).
    ``` {bash}
    cd /usr/203b-data/mimic-iv/core
    zcat admissions.csv.gz | head -10
    ```
The unique patients can be found by counting the number of unique subject_id in the first column. Note that the first row is heading, so the total number of unique patients should be 257366.
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz |
    awk -F ',' '{ print $1 }' | sort | uniq -c | wc -l 
    ```

The possible values taken by `admission_type`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $6 }' | sort | uniq -c
    ```

The possible values taken by `admission_location`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $7 }' | sort | uniq -c
    ```

The possible values taken by `insurance`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $9 }' | sort | uniq -c
    ```

The possible values taken by `marital_status`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $11 }'| sort | uniq -c
    ```

The possible values taken by `languages`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $10 }' | sort | uniq -c
    ```

The possible values taken by `ethics`: 
    ``` {bash}
    zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | 
    awk -F ',' '{ print $12 }' | sort | uniq -c
    ```


## Q3. Who's popular in Price and Prejudice

1. You and your friend just have finished reading *Pride and Prejudice* by Jane Austen. Among the four main characters in the book, Elizabeth, Jane, Lydia, and Darcy, your friend thinks that Darcy was the most mentioned. You, however, are certain it was Elizabeth. Obtain the full text of the novel from <http://www.gutenberg.org/cache/epub/42671/pg42671.txt> and save to your local folder. 
    ```{bash, eval=FALSE}
    curl http://www.gutenberg.org/cache/epub/42671/pg42671.txt > pride_and_prejudice.txt
    ```
Do **not** put this text file `pride_and_prejudice.txt` in Git. Using a `for` loop, how would you tabulate the number of times each of the four characters is mentioned?

**Solution:** We can see from the results that Elizabeth is the most mentioned name, appreared 634 in the book. Darcy came the second, with a number of 417.
    ```{bash eval=FALSE}
    declare -a Names=("Elizabeth" "Jane" "Lydia" "Darcy" )
    for var in "${Names[@]}"
      do
      grep -o $var pride_and_prejudice.txt | wc -l
    done
    ```

2. What's the difference between the following two commands?
    ```{bash}
    echo 'hello, world' > test1.txt
    ```
    and
    ```{bash}
    echo 'hello, world' >> test2.txt
    ```
**Solution:** The `>` command overwrites the test1 file once we run the code, and the `>>`` command appends the test to the test2 file.


3. Using your favorite text editor (e.g., `vi`), type the following and save the file as `middle.sh`:
**Solution:** Done.See below the appended code for .sh file. 
Please kindly note that the vi command is extremely slow for me to execute on the teaching server, probably because of the VPN setting. I have tried the exact same code on my local terminal and it works just fine. So I did not execute theses two commands but only demonstrate them below. The file has been uploaded to the repository.

I first create the file by `vi`, then use `i` bottom to get into inser mode. After modifying the file, I used `wq` to save and quit.
    ```{bash eval=FALSE}
    vi middle.sh
    i
    #!/bin/sh
    # Select lines from the middle of a file.
    # Usage: bash middle.sh filename end_line num_lines
    head -n "$2" "$1" | tail -n "$3"
    wq
    ```
Using `chmod` make the file executable by the owner, and run 
    **Solution:**
    Again this command is not excecuted because of the server issue. I used       700 to make the file executable by the owner only.
    ``` {bash eval=FALSE}
    chmod 700 middle.sh
    ./middle.sh pride_and_prejudice.txt 20 5
    ```
Explain the output. Explain the meaning of `"$1"`, `"$2"`, and `"$3"` in this shell script. Why do we need the first line of the shell script?
**Solution:** 
The `"$1"`, `"$2"`, and `"$3"` are the position parameters of the content     in the file. The first line of the shell script shows the environment         (shell) it is using.
    When I run the code on my local terminal, I got 
    Author: Jane Austen
    Editor: R. W. (Robert William) Chapman
Those are the first and second lines in the objective files, following the    command we put in the middle.sh file. 
    
    
### Q4. More fun with Linux

Try these commands in Bash and interpret the results: `cal`, `cal 2021`, `cal 9 1752` (anything unusual?), `date`, `hostname`, `arch`, `uname -a`, `uptime`, `who am i`, `who`, `w`, `id`, `last | head`, `echo {con,pre}{sent,fer}{s,ed}`, `time sleep 5`, `history | tail`.

**Solution:** 
`cal` displays the calender for current month.
    ```{bash}
    cal
    ```
`cal 2021` displays the calender for the year 2021.
    ```{bash}
    cal 2021
    ```
`cal 9 1752` displays the calender for Sep 1752. Note that day 3-13 are missing. I This is only happening to Sep 1752 because the 11 days were skipped to make uo for lack of leap year adjustment.
    ```{bash}
    cal 9 1752
    ```
`date` shows the current date and time.
    ```{bash}
    date
    ```
`hostname` shows the server we are using.
    ```{bash}
    hostname
    ```
`arch` prints the computer architecture.
    ```{bash}
    arch
    ```
`uname -a` prints the basic system information, and the -a indicates we want all available information to be displayed.
    ```{bash}
    uname -a
    ```
`uptime` shows the information about the time, number of users, how long the system has been running and the system load average.
    ```{bash}
    uptime
    ```
`who am i` returns the username.
    ```{bash}
    who am i
    ```
`who` displays the current running level, also it shows the list of logged in users and their IP addresses.
    ```{bash}
    who
    ```
`w` shows the list of logged in users and other information of them
    ```{bash}
    w
    ```
`id` shows the identity information of the current login account, including UID, gid and so on.
    ```{bash}
    id
    ```
`last|head` show the list of all the users logged in and the time they logged out (shown still logged in if they are currently working on the server).
    ```{bash}
    last | head
    ```
`echo {con,pre}{sent,fer}{s,ed}` is returning exactly what is following the term `echo`.
    ```{bash}
    echo {con,pre}{sent,fer}{s,ed}
    ```
`time sleep 5` the system paused for 5 seconds when excecuting.
    ```{bash}
    time sleep 5
    ```
`history | tail` displays the last 10 commands I ran in the terminal.
    ```{bash}
    history | tail
    ```

**Thank you so much for your time!**







